Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2021-01-26 manuscript/draft_rsfmri.Rmd
correct missspecification of  time points for one subject
enable parallel processing for evaluation of models; script for glass brain; fix wrong time point allocation of subject 61

```{r "Load data", include=FALSE}
final = create_sample_df(group = "all", tp = "all")

@@

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQcfc, echo=FALSE}
res <- read.csv("./report/FDFCcorrelations.csv")
rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')

@@

```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
final=read.csv(file = "./report/all_QA_measures.csv")

@@

```{r FDFC correlation, eval=FALSE, echo=FALSE}
# do not evaluate if saved already
source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
FDFC=calc_qc_fc(final)
write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
```

@@

Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
<!-- ### DMN -->
<!-- #### DMN for three timepoints -->
<!-- cc -->Using the images preprocessed with AROMA+CC, we did find a significant co-activation with the PCC in three clusters associated with average BMI. Omitting mean FD as nuisance covariate yielded similar results but four significant clusters. However, in both, there is significant peak activation within the largest cluster.
<!-- gsr -->In analyses on images that were additionally preprocessed with GSR, no cluster survived statistical thresholds for either ROI.
All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
<!-- ### DMN for two timepoints
BMIONLYIG
Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps). However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
<!-- BMI2TP
Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants. Omitting mean FD did not change these results. -->
<!-- GM mask
Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
 -->

@@

To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models.
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome meausres is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.
<!-- #### Reward -->
As shown in Table \@ref(tab:tableFDmodel), there is a significant co-activation of one cluster with the NAcc. Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
<!-- ### DMN -->
We did not find a any clusters for functional connectivity with the PCC.

@@

<!-- ### DMN -->
<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed. There is a significant association between average BMI and the connecitivity between PCC and three other clusters. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.

>>>>

```{r "Load data", include=FALSE}
final <- create_sample_df(group = "both", tp = "all")

@@

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQcfc, echo=FALSE}
res <- read.csv("report/FDFCcorrelations.csv")
rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')

@@

```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
final=read.csv(file = "report/all_QA_measures.csv")

@@

```{r FDFC correlation, eval=FALSE, echo=FALSE}
# do not evaluate if saved already
source("/data/pt_02161/Analysis/Preprocessing/qa/rs_qa/group_level_QA/QC_FC_correlations/calc_qc_fc_correlations.R")
FDFC=calc_qc_fc(final)
write.csv(FDFC,"report/FDFCcorrelations.csv", row.names = FALSE)
```

@@

Next, we tested for significant effects of the average BMI and the individual change in BMI longitudinally across the three time points.
<!-- ### DMN -->
<!-- #### DMN for three timepoints -->
<!-- cc -->Using the images preprocessed with AROMA+CC, we did find a significant co-activation with the PCC in three clusters associated with average BMI. Omitting mean FD as nuisance covariate yielded similar results but four significant clusters. However, in both, there is significant peak activation within the largest cluster.
<!-- gsr -->In analyses on images that were additionally preprocessed with GSR, no cluster survived statistical thresholds for either ROI.
All results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).
<!-- ### DMN for two timepoints
BMIONLYIG
Upon inclusion of only a subsample entailing participants from the intervention group but not the waiting control group, there was still no effect for average BMI, but a significant effect of change in BMI on co-activation between the seed region NAcc and a voxel ($-18,-42,-9$ [mm])<!-- belonging to a cluster of $k = 2$ with $Z = 4.64$ ($p_{FWEcorr} = 0.049$) (in GSR corrected connectivity maps). However the effect did not survive recalculation with Wild bootstrap ($p_{FWEcorr} = 0.426$)
<!-- BMI2TP
Running the tests on a subsample of only two time points and omitting data from the third did not yield any significant co-activation with the ROI for neither of the preprocessing variants. Omitting mean FD did not change these results. -->
<!-- GM mask
Results obtained from analyses with the grey matter masks were equal to those reported for the brain masks except test statistics for the above reported significant longitudinal effect for the NAcc in the intervention group differed only marginally ($q_{FDRcorr} = 0.048$) and did not survive recalculation with Wild bootstrap either ($q_{FDRcorr} = 0.552$).
In regard to the exploratory expeditions on subsamples, consistent with previous results, no evidence for co-activation was found when using the grey matter mask.
 -->
```{r "BMIagesexfd_PCCcc_avgBMI_deact", out.width = "50%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesexfd_PCCcc_avgBMI_deact.pdf")
```

```{r "BMIagesex_PCCcc_avgBMI_deact", out.width = "85%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIagesex_PCCcc_avgBMI_deact.pdf")
```

@@

To investigate the unique contributions of the average value of and the longitudinal change in mean FD, we constructed two additional models.
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome meausres is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.
<!-- #### Reward -->
As shown in Table \@ref(tab:tableFDmodel), there is a significant co-activation of one cluster with the NAcc. Cluster co-activation appeared irrespective of the preprocessing procedure, differing mainly in cluster size.
<!-- ### DMN -->
We did not find a any clusters for functional connectivity with the PCC.
```{r "FDagesex_NACCcc_avgFD_act", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCcc_avgFD_act.pdf")
```

```{r "FDagesex_NACCgsr_avgFD_act", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/FDagesex_NACCgsr_avgFD_act.pdf")
```

@@

<!-- ### DMN -->
<!-- cc -->The effect of average BMI on connectivity between the PCC and other regions reported previously is confirmed. There is a significant association between average BMI and the connecitivity between PCC and three other clusters. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics of these clusters differ only slightly from the effect found for average BMI when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when thresholding if the images had been GSR corrected.
```{r "BMIFDagesex_NACCcc_avgFD_act", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_avgFD_act.pdf")
```

```{r "BMIFDagesex_NACCcc_cgnBMI_deact", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCcc_cngBMI_deact.pdf")
```

```{r "BMIFDagesex_NACCgsr_avgFD_act", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_avgFD_act.pdf")
```

```{r "BMIFDagesex_NACCgsr_cgnBMI_deact", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_NACCgsr_cngBMI_deact.pdf")
```

```{r "BMIFDagesex_PCCcc_avgBMI_deact", out.width = "75%"}
knitr::include_graphics("/data/pt_02161/Analysis/Project2_resting_state/seed-based/Second_level /code_and_manuscript/report/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
```
